Notebook: Predictors of frog survival following translocation

### 2022-03-19 doc/notebook/notebook_translocation.Rmd
Add notes on structure of dataset and model

### Donor site
* 1-year frog survival is highly variable across sites, but highly repeatable within a site. This indicates the importance of including site_id as a group-level effect. It remains unclear whether also including a group effect (year) that identifies each translocation might also be needed (or feasible given relatively small number of translocations per site). 
* Donor site has a strong effect on frog survival, with 70459 > 70567 > 72996. Explored possible effects of bd_load and frog size as drivers of this pattern. Differences in bd_load between donor sites seems unrelated to survival, but appears closely associated with frog size (see sub-bullets below). Whether frog size is the sole driver or is a results of frog size plus other unmeasured donor site attributes (e.g., degree of resistance against Bd infection) remains to be seen. 
  * Minimal differences between donor sites in median bd_load, but slightly higher at 72996. No obvious effect of bd_load on 1-year frog survival. 
  * Strong effect of frog size on 1-year survival, with frogs /> 55 mm having substantially higher 1-year survival. Frogs />60 mm had by far the highest survival. 
  * The effect of frog size on survival was unrelated to bd_load: No relationship between the two variables. 
  * Frog size differs between donor site: 72996 < 70459 = 70567. This partially mirrors the observed differences in survival between sites.

@@

## Model structure
As originally conceived: survival \~ elevation + shore + snow_t + snow_t1 + year + day + order + donor + sex + length + log(bdload) \| site_id

Based on exploratory analyses: survival \~ elevation + snow_t + snow_t1 + day + order + donor + sex + length + log(bdload) \| site_id + year

Group-level effects: In addition to the inclusion of site_id as a group-level effect, may need to include year of translocation? Given that some sites only received a single translocation, will this cause problems?

Final steps before building model (in analysis Rmd):

* Recode order as 0/1 (instead of 1--4).  
* Recode survival as 0/1 (instead of continuous probality).  
* Recode binary variables as 0/1 (or format as factor): sex (see Gelman reference below).  
* Format some variables as factor: surv, shore, order, donor, sex, site_id, year.  
* Standardize continuous variables: elevation, snow_t, snow_t1, day, length, log(bd_load). Gelman recommends standardizing by dividing by 2 standard deviations. Standardizing allows assessment of importance of predictors by comparing regressions coefficients. Dividing by 2 SD puts continuous variables on approximately the same scale as binary (0/1) variables. This is true only if binary predictors have similar number of 0s and 1s. Sex: female = 428, male = 351; order: 0 = 507, 1 = 272.

@@

@@

To be added soon.

>>>>

### Donor site
* 1-year frog survival is highly variable across sites, but highly repeatable within a site. This indicates the importance of including site_id as a group-level effect. It remains unclear whether also including a group effect (year) that identifies each translocation might also be needed (or feasible given relatively small number of translocations per site). 
* Donor site has a strong effect on frog survival, with 70459 > 70567 > 72996. Explored possible effects of bd_load and frog size as drivers of this pattern. Differences in survival between donor sites seem unrelated to bd_load, but appear closely associated with frog size (see sub-bullets below). Whether frog size is the sole driver or is a results of frog size plus other unmeasured donor site attributes (e.g., degree of resistance against Bd infection) remains to be seen. 
  * Minimal differences between donor sites in median bd_load, but slightly higher at 72996. No obvious effect of bd_load on 1-year frog survival. 
  * Strong effect of frog size on 1-year survival, with frogs /> 55 mm having substantially higher 1-year survival. Frogs />60 mm had by far the highest survival. 
  * The effect of frog size on survival was unrelated to bd_load: No relationship between the two variables. 
  * Frog size differs between donor site: 72996 < 70459 = 70567. This partially mirrors the observed differences in survival between sites.

@@

## Model structure

As originally conceived: survival \~ elevation + shore + snow_t + snow_t1 + year + day + order + donor + sex + length + log(bdload) \| site_id

Based on exploratory analyses: survival \~ elevation + snow_t + snow_t1 + day + order + donor + sex + length + log(bdload) \| site_id

Make pairs plot to identify correlations between predictors (ggpairs from GGally package).

It may not be possible to include both donor and length in a single model because the distribution of frog lengths is quite different across sites, causing dependency between donor and length. As such, these two variables have the potential to compete for the same variation. Need to look at this more closely, perhaps by plotting frog lengths at each donor site. Also, plot length against survival, grouped by donor.  Are these distributions as donor-specific as they apear to be? Because of the importance of distinguishing frog size effects from other effects associated with donor populations (e.g., inherent differences in Bd susceptibility), may need to explore whether it would be possible to compare the fit of two models, each containing one of these two variables.

Also need to disentangle elevation from site_id - currently only one elevation per site_id. Try breaking continuous elevation into categories, with > 2 site_ids per category. Hopefully, that will make elevation variable less dependent on site_id.

Shore and elevation may also be closely associated. Are all sites with shore = 0 at low elevation? If so, may need to only include elevation in model.

Group-level effects: Numerous levels of structure inherent in the data, including that caused by year, site_id, donor, and translocation id. However, most of these are not sufficiently hierarchical to allow inclusion as a group-level effect. For example, 70413 only received frogs collected from 70567. For year, only a small subset of site ids are represented in each year and the subset changes dramaticall from year to year. Donor and translocation id have even sparser matrices with site_id. As such, site_id is the only reasonable group-level effect to consider for inclusion in model. Related to this, general guidance is that 5 or more levels are generally necessary for a grouping variable to be considered for inclusion as a group-level effect. Models with fewer levels may converge but the group-level variable may have little or no effect.

Interactions: General guidance is that interactions between the most important predictors should be included in model.

## Final steps before building model (in analysis Rmd):

* Recode order as 0/1 (instead of 1--4).  
* Recode survival as 0/1 (instead of continuous probability).  
* Recode binary variables as 0/1 indicators: sex (see Gelman reference below).  
* Format some variables as factor: surv, shore, order, donor, sex, site_id, year.  
* Standardize continuous variables: elevation, snow_t, snow_t1, day, length. Possibly standardize log(bd_load) also, but see Gelman for recommendations. Gelman recommends centering by subtracting the mean and standardizing dividing by 2 standard deviations. Centering to a mean of 0 makes interactions more interpretable. Standardizing allows assessment of importance of predictors by comparing regressions coefficients. Dividing by 2 SD puts continuous variables on approximately the same scale as binary (0/1) variables. This is true only if binary predictors have similar number of 0s and 1s. In the current dataset, there is some deviation from 1:1. Sex: female = 428, male = 351; order: 0 = 507, 1 = 272. Binary variables should be centered but not standardized. Variables in need of log-transformation should be transformed and then rescaled, although rescaling will not always be necessary.

@@

## Priors

* For regression coefficients in logistic regression: A recommended weakly informative prior is beta ~ student_t(nu,0,s), where s is chosen to provide weak information on the expected scale, and 3<nu<7.  https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations.  
* Other recommendations in:
  * https://doi.org/10.1111/oik.0598.  
  * Choosing priors in Bayesian ecological models by simulating from the prior predictive distribution.  https://doi.org/10.1002/ecs2.3739

@@

* Figures showing estimated coefficients from models with and without site_id as group-level effect.

## Next steps

* Site_id group-level effect largely accounted for differences between donor sites in frog survival, frog size, etc. To provide another assessment of the effect of frog size on survival, create a second model that includes only donor = 72996. Given that there is considerable variation in frog sizes translocated from this site, this might allow an assessment of the effect of frog size without the group-level effect (because there probably aren't enough sites to which frogs from 72996 were translocated) to make site_id a useful group-level effect.  
* To assess within-site association between survival for the first and second translocated cohorts, build another dataset that includes all of those recipient sites for which > 1 translocation was conducted. Include survival from the first translocation as the response variable and survival from the second translocation as a predictor variable. Other predictors could also be included.
